---
title: "Homework 5"
author: "Anna Schmidt"
date: "2/16/2022"
output: 
  html_document: 
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

___
#### **Question 1**

**Assign to the variable `n_dims` a single random integer between 3 and 10.**

```{r}
n_dims <- sample(3:10, 1)
print(n_dims)
```

**Create a vector of consecutive integers from 1 to n_dims^2.**
```{r}
my_vec <- 1:n_dims^2
```

**Use the sample function to randomly reshuffle these values.**
```{r}
my_vec2 <- sample(my_vec)
```

**create a square matrix with these elements.**
```{r}
m <- matrix(data=my_vec2,nrow=n_dims)
```

**print out the matrix.**
```{r}
print(m)
```

**find a function in r to transpose the matrix.**
```{r}
m_transposed <- t(m)

```

**print it out again and note how it has changed.**
```{r}
print(m_transposed)
```

**calculate the sum and the mean of the elements in the first row and the last row.**
```{r}
# Sum of elements in the first row
firstrow_sum = sum(m_transposed[1,])
print(firstrow_sum)

# Mean of elements in the first row
firstrow_mean = mean(m_transposed[1,])
print(firstrow_mean)

# Sum of elements in the last row
lastrow_sum = sum(m_transposed[n_dims,])
print(lastrow_sum)

# Mean of elements in the last row
lastrow_mean = mean(m_transposed[n_dims,])
print(lastrow_mean)
```

**read about the `eigen()` function and use it on your matrix**
```{r}
m_eigen <- eigen(m_transposed)
print(m_eigen)
```

**look carefully at the elements of `$values` and `$vectors`. What kind of numbers are these?**
```{r}
print(m_eigen$values)
```
The $values are non-integers, and can be negative or positive.
```{r}
print(m_eigen$vectors)
```
The $vectors are non-integers and all fall between -1 and 1. Some are negative and some are positive.

**dig in with the `typeof()` function to figure out their type.**
```{r}
typeof(m_eigen$values)
typeof(m_eigen$vectors)
```

**if have set your code up properly, you should be able to re-run it and create a matrix of different size because `n_dims` will change.**

___
#### **Question 2**

**Create a list with the following named elements:**

**`my_matrix`, which is a 4 x 4 matrix filled with random uniform values**
```{r}
my_vec <- runif(16)
my_matrix <- matrix(data=my_vec, nrow=4)
print(my_matrix)
```

**`my_logical` which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.**
```{r}
my_vec2 <- runif(100)
my_logical <- my_vec2 > 0.5
print(my_logical)
```

**`my_letters`, which is a 26-element vector of all the lower-case letters in random order.**
```{r}
my_letters_ordered <- letters[1:26]
my_letters <- sample(my_letters_ordered)
print(my_letters)
```

**create a list with these**
```{r}
my_list <- list(my_matrix, my_logical, my_letters)
print(my_list)
```

**create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.**
```{r}
my_list_2 <- list(my_matrix[2,2], my_logical[2], my_letters[2])
print(my_list_2)
```

**use the `typeof()` function to confirm the underlying data types of each component in this list**
```{r}
typeof(my_list_2[[1]])
typeof(my_list_2[[2]])
typeof(my_list_2[[3]])
```

**combine the underlying elements from the new list into a single atomic vector with the `c()` function.**
```{r}
my_vector <- c(my_list_2[[1]], my_list_2[[2]], my_list_2[[3]])
print(my_vector)
```

what is the data type of this vector?
```{r}
print(typeof(my_vector))
```

___
#### **Question 3**

**Create a data frame with two variables (= columns) and 26 cases (=rows).**

**call the first variable `my_unis` and fill it with 26 random uniform values from 0 to 10**
```{r}
my_unis <- runif(26, min=0, max=10)
print(my_unis)
```

**call the second variable `my_letters` and fill it with 26 capital letters in random order.**
```{r}
my_letters_ordered <- LETTERS[1:26]
my_letters <- sample(my_letters_ordered)
print(my_letters)
```
  
**create the data frame**
```{r}
df <- data.frame(my_unis, my_letters)
print(df)
```

**for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with `NA`.**
```{r}
df$my_unis[sample(1:26, 4)] <- NA
```

**for the first variable, write a single line of R code to identify which rows have the missing values.**
```{r}
which(is.na(df$my_unis))
```

**for the second variable, sort it in alphabetical order**
```{r}
df[order(df$my_letters), ]
```

**calculate the column mean for the first variable.**
```{r}
mean(df$my_unis, na.rm = TRUE)
```

___
[Back to Home Page](index.html) 
